---
title: "PTSD+AOD NMA"
output:
  html_document:
    df_print: paged
    fig_crop: false
editor_options:
  chunk_output_type: console
---
<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 1800)
```

# Load data and packages
```{r}
df.smd <- readRDS("df.smd.rds")
library("netmeta")
```

# Determine network structure
## PTSD outcome
```{r}
df.smd.ptsd <- netmeta::pairwise(studlab = citation, treat = category, n = n, mean = mean, sd = sd,
                                 data = df.smd[df.smd$outcome == "ptsd",], sm = "SMD")
nc.ptsd <- netconnection(treat1, treat2, studlab, df.smd.ptsd)
print(nc.ptsd)
nc.ptsd$D.matrix
```

## Alcohol outcome (matches with PTSD)
```{r}
df.smd.alcohol <- netmeta::pairwise(studlab = citation, treat = category, n = n, mean = mean, sd = sd,
                                 data = df.smd[df.smd$outcome == "alcohol",], sm = "SMD")
nc.alcohol <- netconnection(treat1, treat2, studlab, df.smd.alcohol)
print(nc.alcohol)
nc.alcohol$D.matrix
```

# PTSD OUTCOME
## SUBNETWORK 1: Behavioral controls
```{r}
sn1.categories <- c("behavioral (aod)",
                    "behavioral (control)",
                    "integrated",
                    "integrated trauma focused",
                    "trauma focused",
                    "integrated and ptsd medication")
#Raw data and pairwise contrasts
df.sn1.ptsd <- netmeta::pairwise(studlab = citation, treat = category, n = n,mean = mean, sd = sd, sm = "SMD", 
                                 data = df.smd[df.smd$outcome == "ptsd" & df.smd[,"category"] %in% sn1.categories,])
#NMA
sn1.ptsd <- netmeta(df.sn1.ptsd, sm = "SMD", fixed = FALSE, random = TRUE,reference.group = "behavioral (control)")
```

### Figure 2A: Network plot
```{r}
#PLOT
#n for node size
df.ptsd.sn1.n <- aggregate(n~category, df.smd[df.smd$outcome == "ptsd" & df.smd[,"category"] %in% sn1.categories,], FUN = sum)
#plot
pdf("Fig2a.pdf", width = 7.5, height = 6)
netgraph(sn1.ptsd, seq = df.ptsd.sn1.n[,"category"], 
         offset = .12, 
         #scale = .8, 
         labels = c("Behavioral\n(AOD)",
                    "Behavioral\n(control)",
                    "Integrated\n(PTSD & AOD)",
                    "Integrated +\nPTSD Medication\n(PTSD & AOD)",
                    "Integrated &\nTrauma Focused\n(PTSD & AOD)",
                    "Trauma Focused\n(PTSD)"),
         points = TRUE, col.points = "gray", 
         cex.points = df.ptsd.sn1.n$n/20, 
         plastic = FALSE, col = "black", thickness = "w.random",
         number.of.studies = TRUE, pos.number.of.studies = .5)
dev.off()
netgraph(sn1.ptsd, seq = df.ptsd.sn1.n[,"category"], 
         offset = .12, 
         #scale = .8, 
         labels = c("Behavioral\n(AOD)",
                    "Behavioral\n(control)",
                    "Integrated\n(PTSD & AOD)",
                    "Integrated + PTSD Medication\n(PTSD & AOD)",
                    "Integrated &\nTrauma Focused\n(PTSD & AOD)",
                    "Trauma Focused\n(PTSD)"),
         points = TRUE, col.points = "gray", 
         cex.points = df.ptsd.sn1.n$n/20, 
         plastic = FALSE, col = "black", thickness = "w.random",
         number.of.studies = TRUE, pos.number.of.studies = .5)
```

### Figure 3A: Results
```{r}
#RESULTS TABLE
netleague(sn1.ptsd, digits = 2)
cat("z statistics, network:")
round(sn1.ptsd$statistic.random, 2)*-1
cat("p-value statistics, network:")
round(sn1.ptsd$pval.random, 3)
cat("z statistics, direct:")
round(sn1.ptsd$statistic.direct.random, 2)
cat("p-value statistics, direct:")
round(sn1.ptsd$pval.direct.random, 3)
```

### Summary
```{r}
summary(sn1.ptsd)
```

## SUBNETWORK 2: Placebo Controls
```{r}
sn2.categories <- c("medication (aod)",
                    "medication (both)",
                    "medication (ptsd)",
                    "placebo (control)",
                    "trauma focused and aod medication",
                    "trauma focused and placebo"
)

#Raw data and pairwise contrasts
df.sn2.ptsd <- netmeta::pairwise(studlab = citation, treat = category, n = n, mean = mean, sd = sd, data = df.smd[df.smd$outcome == "ptsd" & df.smd[,"category"] %in% sn2.categories,], sm = "SMD")
#NMA
sn2.ptsd <- netmeta(df.sn2.ptsd, sm = "SMD", fixed = FALSE, random = TRUE, reference.group = "placebo (control)")
```

### Figure 2B: Network plot
```{r}
#PLOT
#n for node size
df.ptsd.sn2.n <- aggregate(n~category, df.smd[df.smd$outcome == "ptsd" & df.smd[,"category"] %in% sn2.categories,], FUN = sum)
#plot
pdf("Fig2b.pdf", width = 9, height = 6)
netgraph(sn2.ptsd, seq = df.ptsd.sn2.n[,"category"], 
         offset = .12, 
         scale = .8,
         labels = c("Medication\n(AOD)",
                    "Medication\n(PTSD & AOD)",
                    "Medication\n(PTSD)",
                    "Placebo\n(control)",
                    "Trauma Focused\n+ AOD Medication\n(PTSD & AOD)",
                    "Trauma Focused\n+ Placebo\n(PTSD)"),
         points = TRUE, col.points = "gray", 
         cex.points = df.ptsd.sn2.n$n/20, 
         plastic = FALSE, col = "black", thickness = "w.random",
         number.of.studies = TRUE, pos.number.of.studies = .5)
dev.off()
netgraph(sn2.ptsd, seq = df.ptsd.sn2.n[,"category"], 
         offset = .12, 
         #scale = .8, 
         labels = c("Medication\n(AOD)",
                    "Medication\n(PTSD & AOD)",
                    "Medication\n(PTSD)",
                    "Placebo\n(control)",
                    "Trauma Focused\n+ AOD Medication\n(PTSD & AOD)",
                    "Trauma Focused\n+ Placebo\n(PTSD)"),
         points = TRUE, col.points = "gray", 
         cex.points = df.ptsd.sn2.n$n/20, 
         plastic = FALSE, col = "black", thickness = "w.random",
         number.of.studies = TRUE, pos.number.of.studies = .5)
```

### Figure 3B: Results
```{r}
#RESULTS TABLE
netleague(sn2.ptsd, digits = 2)
cat("z statistics, network:")
round(sn2.ptsd$statistic.random, 2)*-1
cat("p-value statistics, network:")
round(sn2.ptsd$pval.random, 3)
cat("z statistics, direct:")
round(sn2.ptsd$statistic.direct.random, 2)
cat("p-value statistics, direct:")
round(sn2.ptsd$pval.direct.random, 3)
```

### Summary
```{r}
summary(sn2.ptsd)
```

# ALCOHOL OUTCOME
## SUBNETWORK 1: Behavioral controls
```{r}
sn1.categories <- c("behavioral (aod)",
                    "behavioral (control)",
                    "integrated",
                    "integrated trauma focused",
                    "trauma focused",
                    "integrated and ptsd medication")
#Raw data and pairwise contrasts
df.sn1.alcohol <- netmeta::pairwise(studlab = citation, treat = category, n = n, mean = mean, sd = sd, data = df.smd[df.smd$outcome == "alcohol" & df.smd[,"category"] %in% sn1.categories,], sm = "SMD")
#NMA
sn1.alcohol <- netmeta(df.sn1.alcohol, sm = "SMD", fixed = FALSE, random = TRUE, reference.group = "behavioral (control)")
```

### Figure 3C: Results
```{r}
#RESULTS TABLE
netleague(sn1.alcohol, digits = 2)
cat("z statistics, network:")
round(sn1.alcohol$statistic.random, 2)*-1
cat("p-value statistics, network:")
round(sn1.alcohol$pval.random, 3)
cat("z statistics, direct:")
round(sn1.alcohol$statistic.direct.random, 2)
cat("p-value statistics, direct:")
round(sn1.alcohol$pval.direct.random, 3)
```

### Summary
```{r}
summary(sn1.alcohol)
```

## SUBNETWORK 2: Placebo Controls
```{r}
sn2.categories <- c("medication (aod)",
                    "medication (both)",
                    "medication (ptsd)",
                    "placebo (control)",
                    "trauma focused and aod medication",
                    "trauma focused and placebo"
)
#Raw data and pairwise contrasts
df.sn2.alcohol <- netmeta::pairwise(studlab = citation, treat = category, n = n, mean = mean, sd = sd, data = df.smd[df.smd$outcome == "alcohol" & df.smd[,"category"] %in% sn2.categories,], sm = "SMD")
#NMA
sn2.alcohol <- netmeta(df.sn2.alcohol, sm = "SMD", fixed = FALSE, random = TRUE, reference.group = "placebo (control)")
```

### Figure 3D: Results
```{r}
#RESULTS TABLE
netleague(sn2.alcohol, digits = 2)
cat("z statistics, network:")
round(sn2.alcohol$statistic.random, 2)*-1
cat("p-value statistics, network:")
round(sn2.alcohol$pval.random, 3)
cat("z statistics, direct:")
round(sn2.alcohol$statistic.direct.random, 2)
cat("p-value statistics, direct:")
round(sn2.alcohol$pval.direct.random, 3)
```

### Summary
```{r}
summary(sn2.alcohol)
```

# Supplement
## Appendix 1. Study-Level Treatment Category Contrasts
### PTSD Outcome
```{r}
df.smd.ptsd[,1:11]
```
### Alcohol Outcome
```{r}
df.smd.alcohol[,1:11]
```

## Appendix 2. Assessments of Small Study Effects
### PTSD, sn1
```{r}
ord.sn1 <- c("behavioral (control)", "behavioral (aod)", "integrated", "trauma focused", "integrated and ptsd medication", "integrated trauma focused")
pdf("plot.funnel.sn1.ptsd.pdf", width = 13, height = 11)
funnel(sn1.ptsd, order = ord.sn1)
dev.off()
metabias(metagen(TE.adj, seTE, data = funnel(sn1.ptsd, order = ord.sn1)))
```

### PTSD, sn2
```{r}
ord.sn2 <- c("placebo (control)", "trauma focused and placebo", "medication (aod)", "medication (ptsd)", "medication (both)", "trauma focused and aod medication")
pdf("plot.funnel.sn2.ptsd.pdf", width = 13, height = 11)
funnel(sn2.ptsd, order = ord.sn2)
dev.off()
metabias(metagen(TE.adj, seTE, data = funnel(sn2.ptsd, order = ord.sn2)))
```

### alcohol, sn1
```{r}
pdf("plot.funnel.sn1.alcohol.pdf", width = 13, height = 11)
funnel(sn1.alcohol, order = ord.sn1)
dev.off()
metabias(metagen(TE.adj, seTE, data = funnel(sn1.alcohol, order = ord.sn1)), method.bias = "Egger")
metabias(metagen(TE.adj, seTE, data = funnel(sn1.alcohol, order = ord.sn1)), method.bias = "Begg")
metabias(metagen(TE.adj, seTE, data = funnel(sn1.alcohol, order = ord.sn1)), method.bias = "Thompson")

```

### alcohol, sn2
```{r}
pdf("plot.funnel.sn2.alcohol.pdf", width = 13, height = 11)
funnel(sn2.alcohol, order = ord.sn2)
dev.off()
metabias(metagen(TE.adj, seTE, data = funnel(sn2.alcohol, order = ord.sn2)))
```

## Appendix 3. Pairwise meta-analyses
In contrast to the NMAs, unadjusted standard errors are used in the calculations and the between-study heterogeneity variance is allowed to differ between comparisons.
### Forest plot, PTSD, sn1
```{r, fig.height= 13.5}
net.pw.sn1.ptsd <- netpairwise(sn1.ptsd)
pdf("plot.forest.sn1.ptsd.pdf", width = 8, height = 13.5)
forest(net.pw.sn1.ptsd, leftcols = c("studlab", "n.e", "n.c", "effect", "ci"), rightcols = FALSE,
       label.right = "Favors Second\nTreatment Category",
label.left = "Favors First\nTreatment Category"
)
dev.off()
forest(net.pw.sn1.ptsd, leftcols = c("studlab", "n.e", "n.c", "effect", "ci"), rightcols = FALSE,
       label.right = "Favors Second\nTreatment Category",
label.left = "Favors First\nTreatment Category"
)
```

### Forest plot, PTSD, sn2
```{r, fig.height = 11}
net.pw.sn2.ptsd <- netpairwise(sn2.ptsd)
pdf("plot.forest.sn2.ptsd.pdf", width = 8, height = 11)
forest(net.pw.sn2.ptsd, leftcols = c("studlab", "n.e", "n.c", "effect", "ci"), rightcols = FALSE,
       label.right = "Favors Second\nTreatment Category",
label.left = "Favors First\nTreatment Category"
)
dev.off()
forest(net.pw.sn2.ptsd, leftcols = c("studlab", "n.e", "n.c", "effect", "ci"), rightcols = FALSE,
       label.right = "Favors Second\nTreatment Category",
label.left = "Favors First\nTreatment Category"
)
```

### Forest plot, alcohol, sn1
```{r, fig.height = 13.5}
net.pw.sn1.alcohol <- netpairwise(sn1.alcohol)
pdf("plot.forest.sn1.alcohol.pdf", width = 8, height = 13.5)
forest(net.pw.sn1.alcohol, leftcols = c("studlab", "n.e", "n.c", "effect", "ci"), rightcols = FALSE,
       label.right = "Favors Second\nTreatment Category",
label.left = "Favors First\nTreatment Category"
)
dev.off()
forest(net.pw.sn1.alcohol, leftcols = c("studlab", "n.e", "n.c", "effect", "ci"), rightcols = FALSE,
       label.right = "Favors Second\nTreatment Category",
label.left = "Favors First\nTreatment Category"
)
```

### Forest plot, alcohol, sn2
```{r, fig.height = 11}
net.pw.sn2.alcohol <- netpairwise(sn2.alcohol)
pdf("plot.forest.sn2.alcohol.pdf", width = 8, height = 11)
forest(net.pw.sn2.alcohol, leftcols = c("studlab", "n.e", "n.c", "effect", "ci"), rightcols = FALSE,
       label.right = "Favors Second\nTreatment Category",
label.left = "Favors First\nTreatment Category"
)
dev.off()
forest(net.pw.sn2.alcohol, leftcols = c("studlab", "n.e", "n.c", "effect", "ci"), rightcols = FALSE,
       label.right = "Favors Second\nTreatment Category",
label.left = "Favors First\nTreatment Category"
)
```

##Appendix 4. Sensitivity analyses without Hien 2015, Ruglass 2017 & Pearson 2019
```{r}
sen1.categories <- c("behavioral (aod)",
                    "behavioral (control)",
                    "integrated",
                    "integrated trauma focused",
                    "trauma focused")
#Raw data and pairwise contrasts
df.sen1.ptsd <- netmeta::pairwise(studlab = citation, treat = category, n = n,mean = mean, sd = sd, sm = "SMD", 
                                 data = df.smd[df.smd$outcome == "ptsd" & df.smd[,"category"] %in% sen1.categories 
                                               & df.smd$citation != "Ruglass 2017" & df.smd$citation != "Pearson 2019",])
#NMA
sen1.ptsd <- netmeta(df.sen1.ptsd, sm = "SMD", fixed = FALSE, random = TRUE,reference.group = "behavioral (control)")
```

### Network plot, SN1 without Hien 2015, Ruglass 2017 & Pearson 2019
```{r}
#PLOT
#n for node size
df.ptsd.sen1.n <- aggregate(n~category, df.smd[df.smd$outcome == "ptsd" & df.smd[,"category"] %in% sen1.categories & df.smd$citation != "Ruglass 2017" & df.smd$citation != "Pearson 2019",], FUN = sum)
#plot
pdf("Figsen1.pdf", width = 7.5, height = 6)
netgraph(sen1.ptsd, seq = df.ptsd.sen1.n[,"category"], 
         offset = .12, 
         #scale = .8, 
         labels = c("Behavioral\n(AOD)",
                    "Behavioral\n(control)",
                    "Integrated\n(PTSD & AOD)",
                    "Integrated &\nTrauma Focused\n(PTSD & AOD)"),
         points = TRUE, col.points = "gray", 
         cex.points = df.ptsd.sen1.n$n/20, 
         plastic = FALSE, col = "black", thickness = "w.random",
         number.of.studies = TRUE, pos.number.of.studies = .5)
dev.off()
netgraph(sen1.ptsd, seq = df.ptsd.sen1.n[,"category"], 
         offset = .12, 
         #scale = .8, 
         labels = c("Behavioral\n(AOD)",
                    "Behavioral\n(control)",
                    "Integrated\n(PTSD & AOD)",
                    "Integrated &\nTrauma Focused\n(PTSD & AOD)"),
         points = TRUE, col.points = "gray", 
         cex.points = df.ptsd.sen1.n$n/20, 
         plastic = FALSE, col = "black", thickness = "w.random",
         number.of.studies = TRUE, pos.number.of.studies = .5)
```

### PTSD Results, SN1 without Hien 2015, Ruglass 2017 & Pearson 2019
```{r}
#RESULTS TABLE
netleague(sen1.ptsd, digits = 2)
cat("z statistics, network:")
round(sen1.ptsd$statistic.random, 2)*-1
cat("p-value statistics, network:")
round(sen1.ptsd$pval.random, 3)
cat("z statistics, direct:")
round(sen1.ptsd$statistic.direct.random, 2)
cat("p-value statistics, direct:")
round(sen1.ptsd$pval.direct.random, 3)
```

### PTSD Summary, SN1 without Hien 2015, Ruglass 2017 & Pearson 2019
```{r}
summary(sen1.ptsd)
netsplit(sen1.ptsd)
```

### Alcohol Results, SN1 without Hien 2015, Ruglass 2017 & Pearson 2019
```{r}
df.sen1.alcohol <- netmeta::pairwise(studlab = citation, treat = category, n = n, mean = mean, sd = sd, data = df.smd[df.smd$outcome == "alcohol" & df.smd[,"category"] %in% sen1.categories & df.smd$citation != "Ruglass 2017" & df.smd$citation != "Pearson 2019",], sm = "SMD")
#NMA
sen1.alcohol <- netmeta(df.sen1.alcohol, sm = "SMD", fixed = FALSE, random = TRUE, reference.group = "behavioral (control)")
#RESULTS TABLE
netleague(sen1.alcohol, digits = 2)
cat("z statistics, network:")
round(sen1.alcohol$statistic.random, 2)*-1
cat("p-value statistics, network:")
round(sen1.alcohol$pval.random, 3)
cat("z statistics, direct:")
round(sen1.alcohol$statistic.direct.random, 2)
cat("p-value statistics, direct:")
round(sen1.alcohol$pval.direct.random, 3)
```

### Alcohol Summary, SN1 without Hien 2015, Ruglass 2017 & Pearson 2019
```{r}
summary(sen1.alcohol)
```

## Appendix 5. Sensitivity analyses without Back 2016 (only study that targeted alcohol or substance use- all others were AUD)
```{r}
sen2.categories <- c("medication (aod)",
                    "medication (both)",
                    "medication (ptsd)",
                    "placebo (control)",
                    "trauma focused and aod medication",
                    "trauma focused and placebo"
)

#Raw data and pairwise contrasts
df.sen2.ptsd <- netmeta::pairwise(studlab = citation, treat = category, n = n, mean = mean, sd = sd, data = df.smd[df.smd$outcome == "ptsd" & df.smd[,"category"] %in% sen2.categories & df.smd$citation != "Back 2016",], sm = "SMD")
#NMA
sen2.ptsd <- netmeta(df.sen2.ptsd, sm = "SMD", fixed = FALSE, random = TRUE, reference.group = "placebo (control)")
```

### Network plot, SN2 without Back 2016
```{r}
#PLOT
#n for node size
df.ptsd.sen2.n <- aggregate(n~category, df.smd[df.smd$outcome == "ptsd" & df.smd[,"category"] %in% sen2.categories  & df.smd$citation != "Back 2016",], FUN = sum)
#plot
pdf("Figsen2.pdf", width = 9, height = 6)
netgraph(sen2.ptsd, seq = df.ptsd.sen2.n[,"category"], 
         offset = .12, 
         scale = .8,
         labels = c("Medication\n(AOD)",
                    "Medication\n(PTSD & AOD)",
                    "Medication\n(PTSD)",
                    "Placebo\n(control)",
                    "Trauma Focused\n+ AOD Medication\n(PTSD & AOD)",
                    "Trauma Focused\n+ Placebo\n(PTSD)"),
         points = TRUE, col.points = "gray", 
         cex.points = df.ptsd.sen2.n$n/20, 
         plastic = FALSE, col = "black", thickness = "w.random",
         number.of.studies = TRUE, pos.number.of.studies = .5)
dev.off()
netgraph(sen2.ptsd, seq = df.ptsd.sen2.n[,"category"], 
         offset = .12, 
         #scale = .8, 
         labels = c("Medication\n(AOD)",
                    "Medication\n(PTSD & AOD)",
                    "Medication\n(PTSD)",
                    "Placebo\n(control)",
                    "Trauma Focused\n+ AOD Medication\n(PTSD & AOD)",
                    "Trauma Focused\n+ Placebo\n(PTSD)"),
         points = TRUE, col.points = "gray", 
         cex.points = df.ptsd.sen2.n$n/20, 
         plastic = FALSE, col = "black", thickness = "w.random",
         number.of.studies = TRUE, pos.number.of.studies = .5)
```

### PTSD Results, SN2 without Back 2016
```{r}
#RESULTS TABLE
netleague(sen2.ptsd, digits = 2)
cat("z statistics, network:")
round(sen2.ptsd$statistic.random, 2)*-1
cat("p-value statistics, network:")
round(sen2.ptsd$pval.random, 3)
cat("z statistics, direct:")
round(sen2.ptsd$statistic.direct.random, 2)
cat("p-value statistics, direct:")
round(sen2.ptsd$pval.direct.random, 3)
```

### PTSD Summary, SN2 without Back 2016
```{r}
summary(sen2.ptsd)
```

### Alcohol Results, SN2 without Back 2016
```{r}
#Raw data and pairwise contrasts
df.sen2.alcohol <- netmeta::pairwise(studlab = citation, treat = category, n = n, mean = mean, sd = sd, data = df.smd[df.smd$outcome == "alcohol" & df.smd[,"category"] %in% sen2.categories & df.smd$citation != "Back 2016",], sm = "SMD")
#NMA
sen2.alcohol <- netmeta(df.sen2.alcohol, sm = "SMD", fixed = FALSE, random = TRUE, reference.group = "placebo (control)")
#RESULTS TABLE
netleague(sen2.alcohol, digits = 2)
cat("z statistics, network:")
round(sen2.alcohol$statistic.random, 2)*-1
cat("p-value statistics, network:")
round(sen2.alcohol$pval.random, 3)
cat("z statistics, direct:")
round(sen2.alcohol$statistic.direct.random, 2)
cat("p-value statistics, direct:")
round(sen2.alcohol$pval.direct.random, 3)
```

### PTSD Summary, SN2 without Back 2016
```{r}
summary(sen2.alcohol)
```

# Session information
```{r}
sessionInfo()
```